\begin{frame}
  \frametitle{Dragon of Loowater}

  \begin{block}{Input}
    \begin{itemize}
      \item list of Dragon heads and their sizes
      \item list of Knights and their sizes
      \item a knight can only defeat a head if he is bigger
    \end{itemize}
  \end{block}

  \begin{exampleblock}{Output}
    \begin{itemize}
      \item What is the minimum cost to defeat the dragon? (knight size)
      \item Or is it impossible?
    \end{itemize}
  \end{exampleblock}
\end{frame}

\begin{frame}
  \frametitle{Stern-Brocot Number}

  \begin{block}{Input}
    \begin{itemize}
    \item Two numbers that make a fraction (5 / 7)
    \end{itemize}
  \end{block}

  \begin{exampleblock}{Output}
    \begin{itemize}
      \item Search a "fraction tree" and print the path.
      \item \structure{Implicit Data Structure!}
      \bigskip

      \item Hint: Try to do it by hand a few times.
    \end{itemize}
  \end{exampleblock}
\end{frame}

\begin{frame}
  \frametitle{Bars}

  \begin{block}{}
    \begin{itemize}
    \item Knapsack problem
    \end{itemize}
  \end{block}

  \begin{itemize}
    \item \alert{Tip:} You need to prune to make it in time!
    \item \alert{Tip:} The bar size could be zero!
  \end{itemize}
  \end{frame}

\begin{frame}
  \frametitle{Rat Attack}

  {\small
    \begin{block}{}
      \begin{itemize}
      \item A 1024 x 1024 matrix has rats in some cells;
      \item A bomb with power d kills rats in a 2d+1 x 2d+1 matrix;
      \item Where is the best place to put the bomb?
      \end{itemize}
    \end{block}
    \begin{itemize}
    \item We need to check all squares in the city, including the borders.
    \item But, if we calculate the rats all the time, this becomes very slow.
      (1024 * 1024 * 100 * 100)
    \item It is possible to {\bf Pre-process} the number of rats killed in each bomb position. (1024 * 1024 + 100 * 100 * 20000)
    \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Simple Equations}

  {\small
  \begin{block}{}
    \begin{itemize}
    \item You are given a set of equations: $x + y + z = A, xyz = B, x^2 + y^2 + z^2 = C$
    \item Given A, B, C, what are the values of x, y, z?
    \end{itemize}
  \end{block}

  \begin{itemize}
  \item You need to define limits for x, y and z, how do you do that?
  \item The equation $x^2 + y^2 + z^2 = C$ might be a good place to start.
  \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Through the Desert}
  {\small
    \begin{block}{}
      \begin{itemize}
      \item Simulate a car going through the desert;
      \item Find the least amount of starting fuel needed to win.
      \end{itemize}
    \end{block}

    \begin{itemize}
    \item You could try to calculate the amount of fuel needed in each
      section of the trip (between gas stations);
    \item Or you could just simulate the trip, and make a binary
      search based on the amount of fuel left;
    \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Zones}
  \begin{block}{Input}
    \begin{itemize}
      \item A total number of \structure{towers} and a desired number of towers.
      \item Each tower has a number of \structure{people}
      \item Some people belong to two or three towers!
    \end{itemize}
  \end{block}

  \begin{exampleblock}{Output}
    \begin{itemize}
      \item Select the towers that maximize the number of people!
    \end{itemize}
  \end{exampleblock}
\end{frame}

\begin{frame}
  \frametitle{Little Bishops}

  Like 8 queens, but with bishops!
\end{frame}
