
\section{Conclusion}
\subsection{Summary}
\begin{frame}
  \frametitle{Class Summary}
  \framesubtitle{What did we learn?}
  \begin{itemize}
  \item Graphs come in a wide variety of types;
  \item Graph problems also have many different types;
  \item Most problems involve small modifications of DFS and BFS;
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Class Summary}
  \frametitle{Hints on solving Problems}


  {\smaller
    \begin{block}{Keep a code Library}
      Graph problems/algorithms use a lot of common code. Store a
      template, and copy/paste as necessary!
      \bigskip
      \begin{itemize}
      \item Structures for keeping visited flags and vertex distances;
      \item Structures for keeping parent and children lists;
      \item Adjacency lists;
      \end{itemize}

  \end{block}
  \begin{block}{Common tricky cases -- Be careful!}
    \begin{itemize}
    \item Graphs with 0 or 1 Vertices;
    \item Unconnected Graphs;
    \item Self loops;
    \item Double edges;
    \end{itemize}
  \end{block}
  }
\end{frame}

\begin{frame}
  \frametitle{Next Week}
  \framesubtitle{More Graphs! Specially path finding}
  \begin{itemize}
  \item Shortest Paths (Single Source and All Pairs);
  \item Network Flow (and related problems);
  \item Graph Matching (bipartite matching, etc) (and related problems);
  \end{itemize}
\end{frame}


\subsection{Homework Discussion}

\begin{frame}
  \frametitle{This Week's Problems}
  \begin{itemize}
  \item Dominator
  \item Forwarding Emails
  \item Ordering
  \item Place the Guards
  \item Doves and Bombs
  \item Come and Go
  \item ACM Contest and Blackout
  \item Ancient Messages
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problem Hints (1)}
  {\smaller
    \begin{block}{Dominator}
      \begin{itemize}
        \item If All paths from 0 to node B pass through node A, then node A {\bf dominates} node B;
        \item For all pair of nodes $i,j$, output ``Y'' if $i$ {\bf dominates} $j$, or ``N'' if not;
      \end{itemize}
    \end{block}

    \bigskip

    The idea of this problem is one of ``reachability'' -- can I reach
    node $j$ if I remove node $i$ from the graph?

    \bigskip

    Note: if $j$ is not connected to ``0'', then \emph{no one dominates j}
  }
\end{frame}

\begin{frame}
  \frametitle{Problem Hints (2)}
  {\smaller
    \begin{block}{Forwarding Emails}
      Every person $i$ sends e-mail only to person $j$.

      \bigskip

      What is the longest email chain?

      \bigskip

      Where does it start?
    \end{block}

    \begin{itemize}
    \item How do you deal with loops?
    \item Time limit is not very large, Try to find an O(n) solution!
    \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Problem Hints (3)}
  {\smaller
    \begin{block}{Ordering}
      Print all possible Orderings of a Direct Acyclic Graph

      \bigskip

      Generalize the DAG ordering algorithm which we discussed in class.
    \end{block}

    \begin{block}{Palace Guards}
      \begin{itemize}
        \item How do you represent the roads and junctions as a Graph?
        \item Find a ``guard-no guard'' assignment to vertices of the
          graph.
        \item First test if a solution is possible!
      \end{itemize}
    \end{block}
  }
\end{frame}

\begin{frame}
  \frametitle{Problem Hints (4)}
  {\smaller
    \begin{block}{Doves and Bombs}
      This problem is about finding ``critical vertices'' in a
      graph. But how do you calculate the ``pigeon value'' of a
      vertex?
    \end{block}

    \begin{block}{Come and Go}
      Straight implementation of ``Strong Connected Components''. Be
      careful with tricky graphs!
    \end{block}
  }
\end{frame}

\begin{frame}
  \frametitle{Problem Hints (5)}
  {\smaller
    \begin{block}{ACM Contest and Blackout}
      Goal: Find the {\bf First} minimum spanning Tree and the {\bf
        Second} minimum spanning Tree
    \end{block}

    \bigskip

    \begin{itemize}
    \item In this class we discussed how to find the Minimum Spanning Tree
    \item How would we find the {\bf second minimum?}
    \item Idea: Maybe if we remove some edges from the graph?
    \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Problem Hints (6)}
  \begin{block}{Ancient Message -- Challenge problem!}
    Count the symbols inside an image -- order does not matter!

    \bigskip

    What is the {\bf Main} difference between the symbols?
  \end{block}

  \begin{itemize}
  \item The shape and size of the symbols is actually not important!
  \item Before you begin programming, discover what is the real
    difference between the symbols.
  \item Hint: The numbers ``1'', ``0'', ``8'' have the same difference.
  \end{itemize}
\end{frame}
