\begin{frame}
  \frametitle{This Week's Problems}

  \begin{itemize}
  \item Immediate Decodability
  \item Caesar Cypher
  \item Ensuring Truth
  \item Smeech
  \item String Partition
  \item Prince and Princess
  \item Power Strings
  \item Life Forms
  \end{itemize}

\end{frame}

\begin{frame}[fragile]
  \frametitle{Problem Discussion (1)}
  {\smaller
    \begin{block}{644 -- Immediate Decodability}
      {\bf Input:} A list of binary codes, each representing
      a different symbol.

      \bigskip
      
      {\bf Output:} Whether any symbol in the list is a
      {\bf prefix} of another symbol.
    \end{block}

\begin{verbatim}
Input:                    Output:
 01                       decodable (no prefixes)
 10
 0001
 00101
 9

 001                      not decodable
 0100                     (string 0 is a prefix of 2)
 00101
 01101
 9
\end{verbatim}

\vfill

How to solve this problem?
  }
\end{frame}

\begin{frame}[fragile]
  \frametitle{Problem Discussion (2)}
  
  {\smaller
  \begin{exampleblock}{554 -- Caesar Cypher}
    A {\bf k-rotation cypher} replaces every symbol $N$ with symbol $N+k$,
    including spaces (which are symbol 0).
    \begin{itemize}
    \item {\bf Input:} A list of correct words, and an \structure{encrypted text}
    \item {\bf Output:} Choose the $k$ which matches the maximum number of words in
      the dictionary, and output the \structure{decrypted text}
    \end{itemize}

    \medskip
    
    Notes about the problem:
    \begin{itemize}
    \item Small input: is only one case, crypotext has only 250 characters
    \item Only letters and spaces
    \item Output requirements (linebreak at 60 characters)
    \end{itemize}
  \end{exampleblock}

\vfill
  
\begin{verbatim}
THIS     DAWN     THAT
THE      ZORRO    OTHER
AT       THING
#
BUUBDLA PSSPABUAEBXO
\end{verbatim}
  
  
  }
\end{frame}

\begin{frame}[fragile]
  \frametitle{Problem Discussion (3)}
  
  {\smaller
    \begin{exampleblock}{11357 -- Ensuring Truth (parsing problem)}
      Parse the formula from the input, and evaluate if
      it has any possible ``TRUE'' evaluation.
\begin{verbatim}
INPUT                     OUTPUT
(a&b&c)|(a&b)|(a)         YES
(x&~x)|(a&~a&b)           NO
\end{verbatim}
    \end{exampleblock}

    \begin{block}{11291 -- Smeech (parsing problem)}
      Parse the formula from the input, and evaluate the expected
      value from the formula.
\begin{verbatim}
INPUT                     OUTPUT
7                         7.00
(.5 3 9)                  3.00
(.7 3 (.5 2 2))           3.80
\end{verbatim}
    \end{block}
  }
\end{frame}

\begin{frame}[fragile]
  \frametitle{Problem Discussion (4)}

  {\small
    \begin{block}{String Partition}
      Break a string of digits into numbers, and find the maximum possible sum.

      Hints:
      \begin{itemize}
      \item The maximum integer value is important
      \item Treat this as a search problem
      \end{itemize}
    \end{block}
\begin{verbatim}
INPUT                OUTPUT
1234554321           1234554321
5432112345           543211239
000                  0
121212121212         2121212124
\end{verbatim}
  }
\end{frame}

\begin{frame}[fragile]
  \frametitle{Problem Discussion (5)}

  {\small
    \begin{block}{Prince and Princess}
      The Prince and the princess make a path through the $n*n$
      grid. Both start at square $1$ and end at square $n*n$, but they
      use different paths.
      \medskip

      You have to eliminate steps from both to make their paths identical.
      \medskip

\begin{verbatim}
Input                 Output
1 7 5 4 8 3 9         4 (size of common path)
1 4 3 5 6 2 8 9       (1,5,8,9)

1 3 4 2 5 8 7 10      5 
1 5 8 9 3 2 7 10      (1,5,8,7,10)
\end{verbatim}
    \end{block}
  }
\end{frame}

\begin{frame}[fragile]
  \frametitle{Problem Discussion (6)}

  {\small
    \begin{exampleblock}{10298 -- Power Strings}
      Imagine that a substring $s^n$ is composed of $n$ repetitions of $s$.
      Given $s^n$, find the maximum possible $n$.

      \bigskip

\begin{verbatim}
INPUT                  OUTPUT
abcd                   1
abcabcabc              3
abcabcabcd             1
\end{verbatim}
    \end{exampleblock}

    \vfill
    
    \begin{itemize}
    \item This is a search problem, what would be the naive algorithm?
    \item Note that the size of $s$ is up to $10^6$. So what is a better algorithm?
    \end{itemize}
    
  }
\end{frame}

\begin{frame}[fragile]
  \frametitle{Problem Discussion (7)}

  {\small
    \begin{block}{Life Forms}
      You are given multiple strings and must find{\bf the largest
        substring shared by more than 1/2 of the strings}.

      \medskip

      This is a generalization of the LCS (longest common substring)
      algorithm.

      \medskip

      For some reason, the time limit is 6.66 seconds!
      
    \end{block}

\begin{verbatim}
INPUT                OUTPUT
abcdefg              bcdefg
bcdefgh              cdefgh
cdefghi
\end{verbatim}
    
  }
\end{frame}


% Immediate Decodability -- Brute force, lexi sorting
% Caesar Cypher -- Try all shifts, output formatting
% Ensuring Truth -- Parser. Description looks scary, but it is a
%                -- very simple problem because
%                -- clauses are only one level deep, and bounded by ``or''s
%                -- replace for a more difficult one for next year
% Smeech         -- Parser, recursive calculation of expected value
%                   (simply weighted avg)

% String Partition -- # DP on total sum --
%                  -- from 0 to i calculate sum of 0-i number plus the best
%                  -- of i+1 to the end of the string
%                  -- backprop with memoization - integer max limit works
%                  -- as a breakpoint

% Prince and Princess -- direct application of LCS

% Power Strings -- String matching: Find s in s + s + s + s

% Life Forms -- Largest Common Substring for n/2 of the substrings

