\begin{frame}
  \frametitle{Problem Discussion -- At a glance}

  \begin{itemize}
  \item Dragon of Loowater -- Greedy Search
  \item Stern Brocot Number -- Binary Search
  \item 8 Queens -- Complete and Pruning
  \item Bars -- Knapsack Problem
  \item Social Constraints -- Permutation Search
  \item Rat Attack -- Correct Representation
  \item Simple Equations -- Input Pruning
  \item Through the Desert -- Binary Search
  \end{itemize}  
\end{frame}

\begin{frame}
  \frametitle{Dragon of Loowater}

  {\small 
  \begin{block}{}
    \begin{itemize}
    \item You are given a list of heads and a list of knights;
    \item A knight must be paired with a head that is smaller than him;
    \item The cost is proportional to the Knight's size;
    \item Find the {\bf minimum cost;}
    \end{itemize}
  \end{block}
  
  \begin{itemize}
  \item Direct Greedy Search;
  \item Who should we try to pair first?
    \begin{itemize}
    \item Smallest Knight?
    \item Smallest Head?
    \item Biggest Knight?
    \item Biggest Head?
    \end{itemize}
  \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Stern-Brocot Number}

  {\small
    \begin{block}{}
      \begin{itemize}
      \item Understand a system to create fractions
      \item Find the place of a certain fraction in the system
      \end{itemize}
    \end{block}
    
    \begin{itemize}
    \item Binary search the fractions {\bf by hand};
    \item Create the fractons {\bf as you go}
      \begin{itemize}
      \item You need the \structure{parents}, \structure{current number} and \structure{target}
      \end{itemize}
    \end{itemize}  
  }
\end{frame}

\begin{frame}
  \frametitle{8 Queens}

  {\small
    \begin{block}{}
      \begin{itemize}
      \item The position of {\bf one queen} is {\bf fixed};
      \item Print {\bf all solutions} given the fixed queen;
      \end{itemize}
    \end{block}

    \begin{itemize}
    \item Permutations are your friend!
    \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Bars}

  {\small
  \begin{block}{}
    \begin{itemize}
    \item Knapsack problem
    \item Find a subset of 20 bars that add to the total size.
    \end{itemize}
  \end{block}

  \begin{itemize}
  \item Worst case: You need to test all $2^{20}$ combinations, for 50 test cases.
  \item Pruning!
    \begin{itemize}
    \item If all bars are bigger, it is impossible;
    \item You just need ONE solution;
    \item etc...
    \end{itemize}
  \item The total size {\bf could be zero!}
  \end{itemize}
  
  }
\end{frame}

\begin{frame}
  \frametitle{Social Constraints}

  {\small
    \begin{block}{}
      \begin{itemize}
      \item $N$ people must sit in a row
      \item Constraint 1: $i$ and $j$ must be {\bf less than} $c$ positions away
      \item Constraint 2: $i$ and $j$ must be {\bf more than} $c$ positions away
      \item {\bf How many} valid arrangements exist?
      \end{itemize}
    \end{block}

    \begin{itemize}
    \item You could try to count how many arrangements are eliminated
      by each constraint... \alert{but that is very complicated...}
    \item $N \leq 8$, so the number of arrangements is $8! = 40.000$.
    \item That is a very small number!
    \item It is okay if you test each arrangement by hand (use next\_permutation)
    \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Rat Attack}

  {\small
    \begin{block}{}
      \begin{itemize}
      \item A 1024 x 1024 matrix has rats in some cells;
      \item A bomb with power d kills rats in a 2d+1 x 2d+1 matrix;
      \item Where is the best place to put the bomb?
      \end{itemize}
    \end{block}
    \begin{itemize}
    \item We need to check all squares in the city, including the borders.
    \item But, if we calculate the rats all the time, this becomes very slow.
      (1024 * 1024 * 100 * 100)
    \item It is possible to {\bf Pre-process} the number of rats killed in each bomb position. (1024 * 1024 + 100 * 100 * 20000)
    \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Simple Equations}

  {\small
  \begin{block}{}
    \begin{itemize}
    \item You are given a set of equations: $x + y + z = A, xyz = B, x^2 + y^2 + z^2 = C$
    \item Given A, B, C, what are the values of x, y, z?
    \end{itemize}
  \end{block}

  \begin{itemize}
  \item You need to define limits for x, y and z, how do you do that?
  \item The equation $x^2 + y^2 + z^2 = C$ might be a good place to start.
  \end{itemize}
  }
\end{frame}

\begin{frame}
  \frametitle{Through the Desert}
  {\small
    \begin{block}{}
      \begin{itemize}
      \item Simulate a car going through the desert;
      \item Find the least amount of starting fuel needed to win.
      \end{itemize}
    \end{block}

    \begin{itemize}
    \item You could try to calculate the amount of fuel needed in each
      section of the trip (between gas stations);
    \item Or you could just simulate the trip, and make a binary
      search based on the amount of fuel left;
    \end{itemize}
  }
\end{frame}
