

\begin{frame}
   \frametitle{Problem Discussion -- At a Glance}
   \begin{itemize}
   \item Wedding Shopping -- Example Problem     
   \item Jill Rides Again -- Subset Sum (1D)
   \item Largest Submatrix -- Subset Sum (2D)
   \item Is Bigger Smarter? -- Longest Increasing Subsequence
   \item Murcia's Skyline -- Longest Increasing Subsequence
   \item Trouble of 13 Dots -- 0-1 Knapsack
   \item Exact Change -- Coin Change
   \item Unidirectional TSP -- Pathfinding
   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problem Hints}

  \begin{itemize}
  \item Wedding Shopping
  \item Jill Rides Again
  \end{itemize}

  \vfill
  
  Discussed during class -- just apply the algorithm!
\end{frame}

\begin{frame}
  \frametitle{Problem Hints 2}

  \begin{block}{Largest Submatrix}
    Find the largest rectangles of ``1'' inside a matrix of 1s and 0s.
  \end{block}

  \bigskip

  Hints:
  \begin{itemize}
  \item You want to do a subset-sum to find the rectangle with biggest
    sum (biggest number of 1).
  \item But how do you avoid adding zeroes?
  \end{itemize}

  \bigskip

  This algorithm is often used as a part of other problems, to
  \alert{calculate valid territory}.
\end{frame}

\begin{frame}
  \frametitle{Problem Hints 3}

  \begin{block}{Is bigger Smarter?}
    Out of a set of elephants, you have to find the largest subset that:
    \begin{itemize}
    \item A - Intelligence is decreasing
    \item B - Weight is increasing
    \end{itemize}
    At the same time.
  \end{block}

  \bigskip

  Hints:
  \begin{itemize}
  \item You don't need to calculate both intelligence and weight at
    the same time.
  \item Organize your data well, and you only need to worry about one.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problem Hints 4}
  \begin{block}{Murcia Skyline}
    Find the longest increasing skyline and the longest decreasing skyline.
  \end{block}

  \bigskip

  Hints
  \begin{itemize}
  \item Standard ``Longest Increasing Subsequence''
  \item But the ``longest'' here is modified by the \emph{building width}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problem Hints 4}
  \begin{block}{Trouble of 13 dots}
    Find the subset of items that:
    \begin{itemize}
    \item Mazimize flavor;
    \item Is inside the price budget
    \end{itemize}
  \end{block}

  \bigskip

  Hints
  \begin{itemize}
  \item This is a 1-0 knapsack problem:
  \item Find the set of items that fills in the knapsack maximizing flavor
  \item Beware that \alert{the knapsack change size if price > 2000!}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problem Hints 5}
  \begin{block}{Exact Change}
    Find the smallest amount of overpay that you can do, with the
    smallest number of coins.
  \end{block}

  \bigskip
  Hints
  \begin{itemize}
  \item Variation of the Coin Change problem discussed in Class;
  \item Calculate all possible changes above the desired value, and find the smallest;
  \item Order by smallest number of coins necessary;
  \item Bottom Up algorithm is probably best;
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problem Hints 6}
  \begin{block}{Unidirectional TSP}
    Find the minimal path from left to right. Up and down are connected!
  \end{block}

  \bigskip

  \begin{itemize}
  \item Very similar to the ``apple robot'' problem;
  \item Note that two paths with the same weight, the smaller index is best!
  \end{itemize}
\end{frame}


