\section{Conclusion}

\subsection{Problem Discussion}

\begin{frame}{Problems for this Week}

  \begin{itemize}
  \item Immediate Decodability
  \item Caesar Cypher
  \item Power Strings
  \item Where's Waldorf
  \item Extend to Palindrome
  \item String Partition
  \item Prince and Princess
  \item Power Strings
  \item Life Forms
  \end{itemize}

\end{frame}

\begin{frame}[fragile]{Immediate Decodability}
  \begin{block}{Outline}
    A set of tokens is decodable if it is {\bf impossible} to write a string that can be parsed in more than one way.\bigskip

    Decodability is detected by checking if a token is a prefix of another.
  \end{block}
\begin{verbatim}
Input:                    Output:
 01                       decodable
 10                       (no string is a prefix)
 0001
 00101

 001                      not decodable
 0100                     (001 is a prefix of 00101)
 00101
 01101
\end{verbatim}
\end{frame}

\begin{frame}[fragile]{Caesar Cypher}
  \begin{exampleblock}{Outline}
    A {\bf k-rotation cypher} replaces every symbol $N$ with symbol $N+k$,
    including spaces (which are symbol 0).
    \begin{itemize}
    \item {\bf Input:} A list of correct words, and an \structure{encrypted text}
    \item {\bf Task:} Find the shift that matches the maximum number of words in the dictionary. Output the \structure{decrypted text}
    \end{itemize}

    \medskip

    Notes about the problem:
    \begin{itemize}
    \item Input: Small, No case, no symbols, spaces
    \item Crypto text may contain words not in dictionary
    \item Output requirements (linebreak at 60 characters)
    \end{itemize}
  \end{exampleblock}

\begin{verbatim}
THIS     DAWN     THAT    || BUUBDLA PSSPABUAEBXO
THE      ZORRO    OTHER   || ATTACK ZORRO AT DAWN
AT       THING            ||
\end{verbatim}
\end{frame}

\begin{frame}[fragile]{Power Strings}
    \begin{exampleblock}{Problem Outline}
      You are given a string $s$, and you must find the smallest string $s'$, so that $s = s'+s'+s'+\ldots = (s')^n$. This is equal to finding $s'$ with maximum $n$.
    \end{exampleblock}
    \bigskip

Example Input and Output:
\begin{verbatim}
INPUT           MINIMUM STRING   N
abcd            abcd             1: abcd
abababab        ab               4: ab + ab + ab + ab
kallakalla      kalla            2: kalla + kalla
abababa         abababa          1: abababa
\end{verbatim}\bigskip

This is a mixture of search and string matching. If your search is not very good, you may face TLE, so write your search carefully.
\end{frame}

\begin{frame}[fragile]{Where is Waldorf?}
  \begin{block}{Problem Outline}
    This the traditional magazine challenge: Find words inside a cube of letters. Pay attention:

    \begin{itemize}
      \item Words can be vertical, horizontal or diagonal;
      \item Words can be backwards;
      \item Search is {\bf not case sensitive}
    \end{itemize}
  \end{block}

\begin{verbatim}
abcDEFGhigg         Words:
hEbkWalDork         Waldorf -- 2 5
FtyAwaldORm         Bambi   -- 2 3
FtsimrLqsrc         Betty   -- 1 2
byoArBeDeyv         Dagbert -- ? ?
Klcbqwikomk
strEBGadhrb
yUiqlxcnBjf
\end{verbatim}
\end{frame}

\begin{frame}[fragile]{Extend to Palindrome}
  \begin{block}{Problem Outline}
    You receive a word as input, and you must add the smallest number of letters at the end to make it a palindrome. Examples:
\begin{verbatim}
  alert:       alertrela
  abcba:       abcba
  aaaalll:     aaaalllaaaa
\end{verbatim}
  \end{block}\bigskip

  Hints:
  \begin{itemize}
    \item Which letters do you add to a word to make it a palindrome;
    \item How do you decide if you add a letter or not?
    \item Can you modify the KLM algorithm to help you make the decision?
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{String Partition}
  \begin{block}{Problem Outline}
    You must break a large string of digits into smaller numbers (max size: 32 bit signed integer), so that the sum of the numbers is the largest.\bigskip

    {\bf Hints:}
    \begin{itemize}
    \item The max number of digits in the string is N=500;
    \item Start with a search on the breaking points;
    \item The total sum can be bigger than signed int;
    \end{itemize}
  \end{block}
\begin{verbatim}
INPUT                           OUTPUT
1234554321                      1234554321
5432112345                      543211239
000                             0
111111111111111111111111111111  3333333333
\end{verbatim}
\end{frame}

\begin{frame}[fragile]{Prince and Princess}
  \begin{block}{Problem outline}
    The Prince and the princess make different paths through the same $n*n$ grid. Both paths start and end at the same square.\bigskip

    Your task is to make both paths identical by eliminating steps, and print the size of the common path.
  \end{block}
\begin{verbatim}
Input                 Output
1 7 5 4 8 3 9         4
1 4 3 5 6 2 8 9       (Common path: 1,5,8,9)

1 3 4 2 5 8 7 10      5
1 5 8 9 3 2 7 10      (Common path: 1,5,8,7,10)
\end{verbatim}
\end{frame}


\begin{frame}[fragile]{Life Forms}
  \begin{block}{Problem Outline}
    Given a set of strings, find the {\bf largest common substring} that is shared by more than half of the strings.
    \bigskip

    Hints:
    \begin{itemize}
      \item Generalization of LCS, but for multiple strings;
      \item If there are multiple substrings of the same size, output all of them;
    \end{itemize}
  \end{block}

\begin{verbatim}
INPUT                OUTPUT
abcdefg              bcdefg
bcdefgh              cdefgh
cdefghi
\end{verbatim}
\end{frame}


% Immediate Decodability -- Brute force, lexi sorting
% Caesar Cypher -- Try all shifts, output formatting

% String Partition -- # DP on total sum --
%                  -- from 0 to i calculate sum of 0-i number plus the best
%                  -- of i+1 to the end of the string
%                  -- backprop with memoization - integer max limit works
%                  -- as a breakpoint

% Prince and Princess -- direct application of LCS

% Power Strings -- String matching: Find s in s + s + s + s

% Life Forms -- Largest Common Substring for n/2 of the substrings
